---
title: "Flood Situation in Southern Malawi"
date: "Last updated `r Sys.Date()`"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    source_code: embed
    vertical_layout: fill
runtime: shiny
resource_files:
- Flood Mapping.Rproj
- flood mapping.R
- .Rhistory
- .Renviron
---

```{r global, echo =FALSE, include=FALSE}
# install required packages
if(!require(devtools)) install.packages("devtools", repos = "http://cran.rstudio.com")
if(!require(leaflet)) install.packages("leaflet", repos = "http://cran.rstudio.com")
if(!require(shiny)) install.packages("shiny", repos = "http://cran.rstudio.com")
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.rstudio.com")
if(!require(leaflet)) install.packages("leaflet", repos = "http://cran.rstudio.com")
if(!require(leafsync)) install.packages("leafsync", repos = "http://cran.rstudio.com")
if(!require(leaflet.extras)) install.packages("leaflet.extras", repos = "http://cran.rstudio.com")
if(!require(shinyWidgets)) install.packages("shinyWidgets", repos = "http://cran.rstudio.com")
if(!require(shinydashboard)) install.packages("shinydashboard", repos = "http://cran.rstudio.com")
if(!require(shinythemes)) install.packages("shinythemes", repos = "http://cran.rstudio.com")
if(!require(geojsonio)) install.packages("geojsonio", repos = "http://cran.rstudio.com")
if(!require(RColorBrewer)) install.packages("RColorBrewer", repos = "http://cran.rstudio.com")
if(!require(sf)) install.packages("sf", repos = "http://cran.rstudio.com")
if(!require(flexdashboard)) install.packages("flexdashboard", repos = "http://cran.rstudio.com")
if(!require(shinyjs)) install.packages("shinyjs", repos = "http://cran.rstudio.com")
if(!require(htmltools)) install.packages("htmltools", repos = "http://cran.rstudio.com")
if(!require(Rcpp)) install.packages("Rcpp",repos = "http://cran.rstudio.com")
if(!require(terra)) devtools::install_version("terra", "1.5-17")
if(!require(raster)) install.packages("raster", repos = c("http://rstudio.org/_packages",
"http://cran.rstudio.com"), dependencies = TRUE)
if(!require(rgdal)) install.packages("rgdal",repos = "http://cran.rstudio.com")
if(!require(rmapshaper)) install.packages("rmapshaper")

dirname <-  '~/R/Flood mapper'
if(!dir.exists(dirname))dir.create(dirname,recursive=TRUE)

# load installed libraries
library(flexdashboard)
library(shiny)
library(shinythemes)
library(shinydashboard)
library(shinyWidgets)
library(leaflet)
library(leaflet.extras)
library(leafsync)
library(sf)
library(tidyverse)
library(RColorBrewer)
library(shinyjs)
library(htmltools)
library(raster)
library(rgdal)
library(rmapshaper)

# identify home directory
normalizePath("~/")

# set mapping colour for layers to ensure consistency
flood_col <- "#0084A8" # blue
wetland_col <- "#002673" # dark blue
settlement_col <- "#4C0073" # purple
river_col <- "#64e3fc" # aqua blue 
cropland_col <- "#70A800" # green


damage_assessment_col <- function(data){
  sapply(data$damage_gra, function(damage_gra){
    if(damage_gra == "Damaged"){
      "red"
    } else if(damage_gra == "Possibly damaged"){
      "orange"
    } else{
      "yellow"
    }
  })
}

# Icon for point of interest
poi_icon <- awesomeIcons(icon = "home",
                         iconColor = "purple",
                         library = "glyphicon",
                         markerColor = "white",
                         squareMarker = TRUE)


# import data sourced from Copernicus Emergency Management Service
flooded_area_chikwawa <- st_read("data/EMSR561_AOI01_DEL_PRODUCT_observedEventA_r1_v1.shp")

wetland_chikwawa <- st_read("data/EMSR561_AOI01_DEL_PRODUCT_hydrographyA_r1_v1.shp")

settlement_chikwawa <- st_read("data/EMSR561_AOI01_DEL_PRODUCT_builtUpA_r1_v1.shp") |> 
  rmapshaper::ms_simplify()

settlement_damage_chikwawa <- st_read("data/EMSR561_AOI01_GRA_PRODUCT_builtUpA_r1_v1.shp")|> 
  rmapshaper::ms_simplify()

chikwawa_damage_col <- colorFactor(palette = c("yellow", "red"),
                                   domain = settlement_damage_chikwawa$damage_gra)

rivers_chikwawa <- st_read("data/EMSR561_AOI01_DEL_PRODUCT_hydrographyL_r1_v1.shp")

poi_chikwawa <- st_read("data/Chikwawa_places.shp") |> 
  st_cast("POINT") # convert multipoint geometry to point 

aoi_chikwawa <- st_read("data/EMSR561_AOI01_DEL_PRODUCT_areaOfInterestA_r1_v1.shp")

cropland_chikwawa <- st_read("data/cropland_chikwawa.shp")

flooded_area_bangula <- st_read("data/EMSR561_AOI02_DEL_PRODUCT_observedEventA_r1_v1.shp")

wetland_bangula <- st_read("data/EMSR561_AOI02_DEL_PRODUCT_hydrographyA_r1_v1.shp")

settlement_bangula <- st_read("data/EMSR561_AOI02_DEL_PRODUCT_builtUpA_r1_v1.shp")|> 
  rmapshaper::ms_simplify()

settlement_damage_bangula <- st_read("data/EMSR561_AOI02_GRA_PRODUCT_builtUpA_r1_v1.shp")|> 
  rmapshaper::ms_simplify()

damage_col <- colorFactor(palette = c("red", "orange", "yellow"),
                          domain = settlement_damage_bangula$damage_gra)

rivers_bangula <- st_read("data/EMSR561_AOI02_DEL_PRODUCT_hydrographyL_r1_v1.shp")

poi_bangula <- st_read("data/Bangula_places.shp")|> 
  st_cast("POINT")

aoi_bangula <- st_read("data/EMSR561_AOI02_DEL_PRODUCT_areaOfInterestA_r1_v1.shp") 

cropland_bangula <- st_read("data/cropland Bangula.shp")

flooded_area_nsanje <- st_read("data/EMSR561_AOI03_DEL_PRODUCT_observedEventA_r1_v1.shp")

wetland_nsanje <- st_read("data/EMSR561_AOI03_DEL_PRODUCT_hydrographyA_r1_v1.shp")

settlement_nsanje <- st_read("data/Nsanje residential areas.shp")

built_up_nsanje <- st_read("data/EMSR561_AOI03_DEL_PRODUCT_builtUpA_r1_v1.shp")|> 
  rmapshaper::ms_simplify() # shapefile is large; it needs simplifying
# object.size(built_up_nsanje)

rivers_nsanje <- st_read("data/EMSR561_AOI03_DEL_PRODUCT_hydrographyL_r1_v1.shp")

aoi_nsanje <- st_read("data/EMSR561_AOI03_DEL_PRODUCT_areaOfInterestA_r1_v1.shp")

# import data sourced from United Nations Institute for Training and Research (unitar)
flood_extent_southernMw <- st_read("data/flood_extent_southernMw.gpkg")

# import southern region administration boundary
southern_region <- st_read("data/southern_region_admn_bnd.gpkg")

# import WorldPop population estimates data
#mw_pop_2020 <- raster::raster("data/mwi_ppp_2020_UNadj.tif")

#mw_pop_2020_clip <- raster::crop(mw_pop_2020, extent(as(southern_region, "Spatial"))) |> 
#  raster::mask(as(southern_region, "Spatial"))
#writeRaster(mw_pop_2020_clip, "data/mw_pop_2020_clip.tif")

mw_pop_2020_clip <- raster("data/mw_pop_2020_clip.tif")

#mw_pop_2020_downsized <- aggregate(mw_pop_2020_clip, fact = 20, 
#                                   fun = mean, expand = TRUE)

pop_col <- colorQuantile("RdYlGn", values(mw_pop_2020_clip), n = 4, 
                         reverse = TRUE, na.color = "transparent")

affected_pop_estimates <- function(pop_raster, flood_extent){
  # extract population values
  affected_pop <- round(raster::extract(pop_raster, flood_extent, 
                                        fun= sum, na.rm = TRUE))
  
  affected_pop_df <- affected_pop %>%
    lapply(unlist) %>%
    lapply(as_tibble) %>%
    bind_rows(., .id = "rowID") %>%
    dplyr::rename(pop = value)
  
  flood_extent$rowID <- 1:nrow(flood_extent)
  
  affected_pop_estimates <- merge(flood_extent, affected_pop_df, by = "rowID")
}

# uncomment these lines if running for the first time. 
#affected_pop_southern <- affected_pop_estimates(mw_pop_2020_clip, flood_extent_southernMw) NB: this may take a while...
#total_pop_affected <- sum(affected_pop_southern$pop) # ~108,674 people
#st_write(affected_pop_southern, "data/estimated_pop_affected.shp")

```

## Sidebar {.sidebar}

#### Select flooded area to display

```{r}
# # create a vector of names of flooded areas to choose from
# flooded_area_choices <- c("Chikwawa", "Bangula", "Nsanje")
# 
# # create a drop down menu of flooded area
# selectInput("flooded_area_choice", label = "Display map of floods in:",
#             choices = flooded_area_choices, selected = flooded_area_choices[[1]])

pickerInput("flooded_areas", label = "Display map of floods in:",
            choices = list("Chikwawa", "Bangula", "Nsanje", "Southern region"))

tags$img(src = "legend.png", height = 85, width = 200)

```


##### This dashboard has been prepared to report the riverine flood situation in southern Malawi after the passage of tropical storm Ana on 25 January 2022. The maps show before and after the flood event and are based on based on the [Copernicus Emergency Management Service](https://emergency.copernicus.eu/mapping/list-of-components/EMSR561) and [United Nations Institute for Training and Research](https://www.unitar.org/maps/map/3457). The flood delineation layer has been derived from post-event satellite images (captured on 26/01/2022 and 02/02/2022) using a semi-automatic approach. Details about the tropical storm Ana can be found here: [Africa CDC](https://africacdc.org/news-item/tropical-storm-ana-hits-hard-five-countries-in-the-southern-africa-region/). Population data: [WorldPop](https://www.worldpop.org/geodata/summary?id=27401) (2020) 

##### [Clinton Nkolokosa](https://github.com/ClintyNkolokosa/Flood-Mapper-ShinyApp)

## Column {data-width=400}
### **Flood situation**. Please allow a few moments for the maps to appear. Pan and zoom the map as needed to show more details 
```{r}

renderLeaflet({
  
  map_title <- tags$style(HTML(".leaflet-control.map-title{
                               transform:translate(-20%, 20%);
                               position: fixed!important;
                               top: 15%;
                               left: 26%;
                               text-align: left;
                               padding-left: 10px;
                               padding-right: 10px;
                               background: rgba(255, 255, 2555, 0.75);
                               font-weight: bold;
                               font-size: 15px
                               }"))
  
  leafletitle <- tags$div(map_title, HTML(paste0(reactive_flooded_area()),"<br>",
                                          paste0(reactive_affected_pop())))

  if(input$flooded_areas == "Chikwawa"){
    return(
  # create flood extent map
     leaflet(options = leafletOptions(zoomControl = FALSE)) |>
      htmlwidgets::onRender("function(el, x){
                            L.control.zoom({position: 'bottomright'}).addTo(this)}") |> 
      setView(lng = 34.82414, lat = -16.04908, zoom = 12) |> 
      addProviderTiles(providers$Thunderforest.OpenCycleMap,
                       options = providerTileOptions(apikey = Sys.getenv("APIKEY")),
                       group = "Thunderforest.OpenCycleMap") |> 
      addProviderTiles(providers$Esri.WorldImagery,
                       group = "Esri.WorldImagery") |> 
      addPolylines(data = rivers_chikwawa,
                   weight = 1,
                   smoothFactor = 0.1,
                   color = river_col,
                   opacity = 0.5,
                   group = "River") |> 
     addPolygons(data = cropland_chikwawa,
                 stroke = FALSE,
                 smoothFactor = 0.2,
                 fillOpacity = 0.2,
                 fillColor = cropland_col,
                 group = "Cropland") |> 
     addPolygons(data = settlement_chikwawa,
                 weight = 1,
                 opacity = 0.5,
                 color = "#4C0073",
                 smoothFactor = 0.2,
                 fillOpacity = 0.5,
                 fillColor = settlement_col,
                 group = "Settlement") |> 
      addPolygons(data = settlement_damage_chikwawa,
                  color = ~chikwawa_damage_col(damage_gra),
                  smoothFactor = 0.1,
                  weight = 1,
                  opacity = 0.5,
                  fillOpacity = 0.6,
                  fillColor = ~chikwawa_damage_col(damage_gra),
                  label = lapply(settlement_damage_chikwawa$damage_gra, HTML),
                  group = "Damage") |> 
      addPolygons(data = wetland_chikwawa,
                  stroke = FALSE,
                  smoothFactor = 0.2,
                  weight = 1,
                  color = wetland_col,
                  fillOpacity = 0.5,
                  group = "Pre flood") |>
     addPolygons(data = flooded_area_chikwawa,
                 stroke = FALSE,
                 smoothFactor = 0.2,
                 fillOpacity = 0.5,
                 fillColor = flood_col,
                 label = lapply(flooded_area_chikwawa$notation, HTML),
                 group = "Post flood") |> 
     addPolygons(data = aoi_chikwawa,
                 weight = 1.5,
                 smoothFactor = 0.2,
                 color = "#A80000",
                 fillOpacity = 0,
                 group = "Area of interest") |> 
     addCircleMarkers(data = poi_chikwawa,
                      weight = 1,
                      radius = 5,
                      fillOpacity = 0.4,
                      color = "#4d004b",
                      label = poi_chikwawa$name,
                      labelOptions = labelOptions(
                         style = list("font-weight" = "bold", "color" = " #581845"),
                      textsize = "15px", direction = "auto"))|> 
      addLayersControl(
        overlayGroups = c("Post flood", "Pre flood", "Damage"),
        baseGroups = c("Esri.WorldImagery", "Thunderforest.OpenCycleMap"),
        options = layersControlOptions(collapsed = FALSE)) |> 
      addMiniMap(position = "bottomleft", tiles = (providers$CartoDB.Positron),
                 width = 120, height = 120, collapsedWidth = 10, collapsedHeight =10,
                 zoomLevelOffset = -6, zoomAnimation = TRUE)) |>  
      addControl(leafletitle, position = "topleft", className = "map-title")|> 
      addMeasure(primaryLengthUnit = "kilometers",
                 primaryAreaUnit = "hectares",
                 secondaryAreaUnit = "sqmeters",
                 activeColor = "red",
                 position = "bottomright",
                 completedColor = "#C8F2BE")
  }
  else if(input$flooded_areas == "Bangula"){
    return(
      withProgress(message = 'Chonde dikirani...',
                    value = 1/5,{
      leaflet(options = leafletOptions(zoomControl = FALSE)) |>
      htmlwidgets::onRender("function(el, x){
                            L.control.zoom({position: 'bottomright'}).addTo(this)}") |> 
      setView(lng = 35.09670, lat = -16.56998, zoom = 12) |> # map centered on Bangula and zoomed in a bit                
      addProviderTiles(providers$Thunderforest.OpenCycleMap,
                       options = providerTileOptions(apikey = Sys.getenv("APIKEY")),
                       group = "Thunderforest.OpenCycleMap") |> 
      addProviderTiles(providers$Esri.WorldImagery,
                       group = "Esri.WorldImagery") |> 
      addPolylines(data = rivers_bangula,
                   weight = 1,
                   smoothFactor = 0.1,
                   color = river_col,
                   opacity = 0.5,
                   group = "River") |> 
     addPolygons(data = cropland_bangula,
                 stroke = FALSE,
                 smoothFactor = 0.2,
                 fillOpacity = 0.2,
                 fillColor = cropland_col,
                 group = "Cropland") |> 
     addPolygons(data = settlement_bangula,
                 weight = 0.8,
                 opacity = 0.5,
                 color = "#4C0073",
                 smoothFactor = 0.2,
                 fillOpacity = 0.5,
                 fillColor = settlement_col,
                 group = "Settlement") |> 
      addPolygons(data = settlement_damage_bangula,
                  color = ~damage_col(damage_gra),
                  smoothFactor = 0.1,
                  weight = 1,
                  opacity = 0.5,
                  fillOpacity = 0.6,
                  fillColor =~damage_col(damage_gra),
                  label = lapply(settlement_damage_bangula$damage_gra, HTML),
                  group = "Damage") |> 
      addPolygons(data = wetland_bangula,
                  stroke = FALSE,
                  smoothFactor = 0.2,
                  weight = 1,
                  color = wetland_col,
                  fillOpacity = 0.5,
                  group = "Pre flood") |>
     addPolygons(data = flooded_area_bangula,
                 stroke = FALSE,
                 smoothFactor = 0.2,
                 fillOpacity = 0.5,
                 fillColor = flood_col,
                 label = lapply(flooded_area_bangula$notation, HTML),
                 group = "Post flood") |> 
     addPolygons(data = aoi_bangula,
                 stroke = 1,
                 weight = 1.5,
                 smoothFactor = 0.2,
                 color = "#A80000",
                 fillOpacity = 0,
                 group = "Area of interest") |> 
     addCircleMarkers(data = poi_bangula,
                      weight = 1,
                      radius = 5,
                      fillOpacity = 0.4,
                      color = "#4d004b",
                      label = poi_bangula$name,
                      labelOptions = labelOptions(
                         style = list("font-weight" = "bold", "color" = "#581845"),
                      textsize = "15px", direction = "auto"),
                      group = "Village")|> 
      addLayersControl(
        overlayGroups = c("Post flood", "Pre flood", "Damage"),
        baseGroups = c("Esri.WorldImagery", "Thunderforest.OpenCycleMap"),
        options = layersControlOptions(collapsed = FALSE)) |> 
      addMiniMap(position = "bottomleft", tiles = (providers$CartoDB.Positron),
                 width = 120, height = 120, collapsedWidth = 10, collapsedHeight =10,
                 zoomLevelOffset = -6, zoomAnimation = TRUE)}) |> 
      addControl(leafletitle, position = "topleft", className = "map-title")|> 
      addMeasure(primaryLengthUnit = "kilometers",
                 primaryAreaUnit = "hectares",
                 secondaryAreaUnit = "sqmeters",
                 activeColor = "red",
                 position = "bottomright",
                 completedColor = "#C8F2BE") #|> 
      #addResetMapButton()
    )}
  else if (input$flooded_areas == "Nsanje"){
    return(
       withProgress(message = 'Chonde dikirani...',
                    value = 1/5,{
      leaflet(options = leafletOptions(zoomControl = FALSE)) |> 
      htmlwidgets::onRender("function(el, x){
                            L.control.zoom({position: 'bottomright'}).addTo(this)}") |>                  
      setView(lng = 35.23918, lat = -16.95804, zoom = 12) |>                            
      addProviderTiles(providers$Thunderforest.OpenCycleMap,
                       options = providerTileOptions(apikey = Sys.getenv("APIKEY")),
                       group = "Thunderforest.OpenCycleMap") |> 
      addProviderTiles(providers$Esri.WorldImagery,
                       group = "Esri.WorldImagery") |> 
      addPolylines(data = rivers_nsanje,
                   weight = 1,
                   smoothFactor = 0.1,
                   color = river_col,
                   opacity = 0.5,
                   group = "River") |> 
     addPolygons(data = built_up_nsanje,
                 stroke = 0.3,
                 weight = 0.5,
                 color = "#4C0073",
                 smoothFactor = 5,
                 fillOpacity = 0.6,
                 fillColor = settlement_col,
                 group = "Built up area") |> 
     addPolygons(data = settlement_nsanje,
                 stroke = 1,
                 weight = 1,
                 opacity = 0.5,
                 color = "#4C0073",
                 smoothFactor = 0.2,
                 fillOpacity = 0.5,
                 fillColor = settlement_col,
                 group = "Settlement") |> 
      addPolygons(data = wetland_nsanje,
                  stroke = 0.5,
                  smoothFactor = 0.2,
                  weight = 1,
                  color = wetland_col,
                  fillOpacity = 0.5,
                  group = "Pre flood") |>
     addPolygons(data = flooded_area_nsanje,
                 stroke = FALSE,
                 smoothFactor = 0.2,
                 fillOpacity = 0.5,
                 fillColor = flood_col,
                 group = "Post flood") |> 
     addPolygons(data = aoi_nsanje,
                 stroke = 1,
                 weight = 1.5,
                 smoothFactor = 0.2,
                 color = "#A80000",
                 fillOpacity = 0,
                 group = "Area of interest")|> 
      addLayersControl(
        overlayGroups = c("Post flood", "Pre flood"),
        baseGroups = c("Esri.WorldImagery", "Thunderforest.OpenCycleMap"),
        options = layersControlOptions(collapsed = FALSE)) |> 
      addMiniMap(position = "bottomleft", tiles = (providers$CartoDB.Positron),
                 width = 120, height = 120, collapsedWidth = 10, collapsedHeight =10,
                 zoomLevelOffset = -6, zoomAnimation = TRUE)})|>  
      addControl(leafletitle, position = "topleft", className = "map-title")|> 
      addMeasure(primaryLengthUnit = "kilometers",
                 primaryAreaUnit = "hectares",
                 secondaryAreaUnit = "sqmeters",
                 activeColor = "red",
                 position = "bottomright",
                 completedColor = "#C8F2BE")#|> 
      #addResetMapButton()
    )
  } 
  else {
    return(
       withProgress(message = 'Chonde dikirani...',
                    value = 1/5,{
      leaflet(options = leafletOptions(zoomControl = FALSE)) |> 
      htmlwidgets::onRender("function(el, x){
                            L.control.zoom({position: 'bottomright'}).addTo(this)}") |>
      setView(lng = 35.0554, lat = -15.8417, zoom = 8) |>                            
      addProviderTiles(providers$Thunderforest.OpenCycleMap,
                       options = providerTileOptions(apikey = Sys.getenv("APIKEY")),
                       group = "Thunderforest.OpenCycleMap") |> 
      addProviderTiles(providers$Esri.WorldImagery,
                       group = "Esri.WorldImagery") |>
      addPolygons(data = flood_extent_southernMw,
                  stroke = 0.7,
                  weight = 0.8,
                  opacity = 0.5,
                  color = flood_col,
                  smoothFactor = 0.2,
                  fillOpacity = 0.8,
                  fillColor = flood_col,
                  group = "Flood extent") |> 
      addRasterImage(mw_pop_2020_clip,
                     colors = pop_col,
                     opacity = 0.6,
                     group = "Population") |> 
      addPolygons(data = southern_region,
                  stroke = 1,
                  color = "white",
                  opacity = 0.5,
                  weight = 0.8,
                  fill = TRUE,
                  fillOpacity = 0.1,
                  label = lapply(southern_region$DISTRICT, HTML),
                  popup = ~southern_region$DISTRICT,
                  # add hover effect
                  highlightOptions = highlightOptions(
                    color = "#E2068A",
                    weight = 1.5,
                    bringToFront = TRUE,
                    fillOpacity = 0.5),
                  group = "Southern region") |> 
      addLayersControl(
        overlayGroups = c("Flood extent", "Southern region", "Population"),
        baseGroups = c("Esri.WorldImagery", "Thunderforest.OpenCycleMap"),
        options = layersControlOptions(collapsed = FALSE)) |> 
      addLegend(values = 1:1300000,
                pal = colorQuantile("RdYlGn", 1:1300000, reverse = TRUE),
                title = "Estimated population",
                labFormat = function(type, cuts, p){
                  n = length(cuts)
                  p = paste0(round(p*100), '%')
                  cuts = paste0(prettyNum(cuts[-n], big.mark = ","), " - ",
                                prettyNum(cuts[-1], big.mark = ","))
                  paste0('<span title ="', p[-n], " - ", p[-1], '">', cuts,'</span>')
                },
                opacity = 0.8) |> 
     addMiniMap(position = "bottomleft", tiles = (providers$CartoDB.Positron),
                width = 120, height = 120, collapsedWidth = 10, collapsedHeight =10,
                zoomLevelOffset = -6, zoomAnimation = TRUE)}) |>  
      addControl(leafletitle, position = "topleft", className = "map-title") |> 
      addMeasure(primaryLengthUnit = "kilometers",
                 primaryAreaUnit = "hectares",
                 secondaryAreaUnit = "sqmeters",
                 activeColor = "red",
                 position = "bottomright",
                 completedColor = "#C8F2BE")#|> 
      #addResetMapButton()
    )}

})

 reactive_flood_extent <- reactive({
    if(input$flooded_areas == "Southern region") {
      580
    } else if(input$flooded_areas == "Bangula"){
      33.7
    } else if(input$flooded_areas == "Nsanje"){
      62.4
    } else{
      return(29.7)
    }
  })
  
  reactive_flooded_area <- renderText({
    paste0("Total flooded area: ", reactive_flood_extent()[[1]], "sq.km")
  })

   textOutput("reactive_flooded_area")
   
   
    reactive_affected_people <- reactive({
    if(input$flooded_areas == "Southern region") {
      30500
    } else if(input$flooded_areas == "Bangula"){
      6291
    } else if(input$flooded_areas == "Nsanje"){
      8741
    } else{
      return(2275)
    }
  })
   
   reactive_affected_pop <- renderText({
     paste0("Estimated population affected: ", prettyNum(reactive_affected_people()[[1]], 
                                                         big.mark = ","))
   })
   
   textOutput("reactive_affected_pop")

```


